---
title: "hw4_submission"
format: html
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(purrr)
```

```{r}
# read in data
clim_data <- read.table(here("Data", "clim.txt"),
                        header = TRUE, 
                        sep = " ")
source(here("Assignments/henry_assignment_4/almond_yield.R"))
source(here("Assignments/henry_assignment_4/revenue_npv.R"))
source(here("Assignments/henry_assignment_4/cost_npv.R"))
source(here("Assignments/henry_assignment_4/almond_profit.R"))
```

```{r}
# Get almond yield df
almond_yield <- crop_yield(
  clim_data = clim_data,
  month = month,
  year = year,
  min_temp = tmin_c,
  precip = precip
)

```


```{r}
# Get almond profit df
almond_profit_df <- almond_profit(
  yield_df = almond_yield,
  base_price = 500, # Price per ton of almonds, scales with inflation
  base_cost = 1000, # Fixed cost - does not change with yield, does change with inflation
  inflation_rate = 0.03,
  discount_rate = 0.12
)
```

# Sensitivity test
Boxplots 

1. Fixed cost
```{r}
# make the varied fixed cost values
var_base_cost <- tibble(var_base_cost = seq(from = 500, to = 5000, by = 50))

# add the varied fixed cost values to the almond yield data frame
fixed_cost_senstest <- expand_grid(almond_yield, var_base_cost)

```
The fixed costs are the only input we want to vary. Run the `almond_profit` function with the new dataframe.
```{r}
# Get almond profit df
almond_profit_senstest_fc <- almond_profit(
  yield_df = fixed_cost_senstest,
  base_price = 500, # Price per ton of almonds, scales with inflation
  base_cost = fixed_cost_senstest$var_base_cost, # Fixed cost - does not change with yield, does change with inflation
  inflation_rate = 0.03,
  discount_rate = 0.12
)
```

Let's plot them in a boxplot to analyze how much variation there is in each year based on the variance of fixed costs. 
```{r}
almond_profit_senstest_fc |> 
  mutate(year = as.factor(year)) |> 
ggplot(aes(x = year, y = profit_npv, group = year)) +
  geom_boxplot() +
  labs(title = "NPV Profit - Fixed cost sensitivity test",
       y = "NPV Profit ($)") +
  theme_minimal()
  
```
Based on the Almond profit model and the sensitivity test run above, changing fixed profit values does not significantly effect NPV profit. 


2. Inflation rate 

```{r}
# make the varied price values
var_base_price <- tibble(var_base_price = seq(from = 100, to = 800, by = 50))

# add the varied price values to the almond yield data frame
price_senstest <- expand_grid(almond_yield, var_base_price)

```
The price are the only input we want to vary. Run the `almond_profit` function with the new dataframe.
```{r}
# Get almond profit df
almond_profit_senstest_p <- almond_profit(
  yield_df = price_senstest,
  base_price = price_senstest$var_base_price, # Price per ton of almonds, scales with inflation
  base_cost = 1000, # Fixed cost - does not change with yield, does change with inflation
  inflation_rate = 0.03,
  discount_rate = 0.12
)
```

Let's plot them in a boxplot to analyze how much variation there is in each year based on the variance of price. 
```{r}
almond_profit_senstest_p |> 
  mutate(year = as.factor(year)) |> 
ggplot(aes(x = year, y = profit_npv, group = year)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "NPV Profit - Inflation rate sensitivity test",
       y = "NPV Profit ($)")
  
```
