---
title: "assignment6_forestgrowth"
format: html
date: February 24, 2026
name: Sofia Rodas
message: false
warning: false
---

# Sensitivity Analysis of a Differential Equation
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(deSolve)
library(here)
library(sensitivity)
set.seed(98)
```

## Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C,
and using the following parameters
```{r}
# source our function
source(here("R/forest_growth.R"))

# parms value setting
K <- 250 
r <- 0.01
g <- 2 

forest_growth <- cbind.data.frame(r = r, K = K, g = g) 

# initial forest setting
initial_forest <- 10

# create a time sequence
tm <- seq(from = 1, to = 300)

# watch it grow
forest <- ode(y = initial_forest, times = tm, forestgrowth, parms = forest_growth)

# rename the column names
colnames(forest) <- c("time", "dC")
```

## Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no
uncertainty)
```{r}
# graph the results of forest growth
ggplot(as.data.frame(forest), aes(time, dC)) +
  geom_line(color = "hotpink") + 
  theme_minimal() +
  labs(title = "Forest growth change over time",
        y = "Change in growth (kgC)")
```

## Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the
estimated maximum forest size (e.g maximum of 300 years) varies with these parameters
```{r}
# set the number of parameters
np <- 2000

# Assume parameters are normally distributed
# sample K with mean of 200, standard devidation of  50
K <- rnorm(mean = 250, sd = 25, n = np)

# sample r with mean of 0.01 and standard deviation of 0.001
r <- rnorm(mean = 0.01, sd = 0.001, n = np)

# sample g with mean of 2 and standard deviation of 0.2
g <- rnorm(mean = 2, sd = 0.2, n = np)

# sample thresh with mean of 50 and standard deviation of 5
thresh <- rnorm(mean = 50, sd = 5, n = np)

X1 <- cbind.data.frame(r = r, K = K, g = g, thresh = thresh)

# Assume parameters are normally distributed
# sample K with mean of 200, standard devidation of  50
K2 <- rnorm(mean = 250, sd = 25, n = np)

# sample r with mean of 0.01 and standard deviation of 0.001
r2 <- rnorm(mean = 0.01, sd = 0.001, n = np)

# sample g with mean of 2 and standard deviation of 0.2
g2 <- rnorm(mean = 2, sd = 0.2, n = np)

# sample thresh with mean of 50 and standard deviation of 5
thresh2 <- rnorm(mean = 50, sd = 5, n = np)

X2 <- cbind.data.frame(r = r2, K = K2, g = g2, thresh = thresh2)

# fix any negative values and they are not meaningful
X1 <- X1 |>  
  map_df(pmax, 0.0)

X2 <- X2 |>  
  map_df(pmax, 0.0)

# create our sobel object and get sets of parameters for running the model
sens_forest_size <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# our parameter sets are
head(sens_forest_size$X)

# lets add names
colnames(sens_forest_size$X) <- c("r", "K", "g", "thresh")

head(sens_forest_size$X)
```
```{r}
# turn computing our metrics into a function
compute_pop_metrics <- function(result) {
  # first add maximum population
  maxgrowth <- max(result$C)

  # return results
  return(list(maxgrowth = maxgrowth))
}

# test that the metrics computing function works works
# initial population
initial_forest <- 10

# run for year = 300
simtimes <-  seq(from = 1, to = 300)

# use the first parameter set for the ode function 
parmtest <- list(r = sens_forest_size$X[1,"r"], 
                 K = sens_forest_size$X[1,"K"], 
                 g = sens_forest_size$X[1,"g"],
                 thresh = sens_forest_size$X[1, "thresh"])

# run ode
result <- ode(y = initial_forest, 
              times = simtimes, 
              func = forestgrowth, 
              parms = parmtest)

# turn ode results into a labelled data frame 
result <- data.frame(time=result[,1], C=result[,2])
```

### Make a wrapper function to run a Sobol test 
```{r}
# define a wrapper function to do everything we need - run solver and compute metrics - and send back results 
p_wrapper <- function(r, K, g, thresh, initial_forest, simtimes, odefunc, metricfunc) {
  
  # create parameter list
  parms <- list(r = r, K = K, g = g, thresh = thresh)
  
  # run ode
  result <- ode(y = initial_forest, times = simtimes, func = odefunc, parms = parms, method="euler")
  
  # results as a data frame
  result <- data.frame(time=result[,1], C=result[,2])
  
  # compute metrics
   metrics <- metricfunc(as.data.frame(result))
   
  return(metrics)
}
```



```{r}
# test with specific r and K values (r=0.01, K=250)
# set the threshold 
metricpars = list(thresh=50)
result <- p_wrapper(r = 0.01, 
          K = 250, 
          g = 2, 
          thresh = 50, 
          initial_forest = 10, 
          simtimes = simtimes,
          odefunc = forestgrowth,
          metricfunc = compute_pop_metrics
          )

result
```

```{r}
# try with first row of our Sobol parameters
p_wrapper(r = sens_forest_size$X[1,"r"],
          K = sens_forest_size$X[1,"K"],
          g = sens_forest_size$X[1,"g"],
          thresh = sens_forest_size$X[1,"thresh"],
          initial_forest = initial_forest,
          simtimes = simtimes,
          odefunc = forestgrowth,
          metricfunc = compute_pop_metrics)
```


```{r}
# run the wrapper function over the 
allresults <- as.data.frame(sens_forest_size$X)  |>  pmap(p_wrapper,
                                                          initial_forest = initial_forest,
                                                          simtimes = simtimes,
                                                          odefunc = forestgrowth,
                                                          metricfunc = compute_pop_metrics)

# extract  results from pmap into a data frame
allres <- allresults |>  map_dfr(`[`, c("maxgrowth"))
```

## Graph the results of the sensitivity analysis as a box plot of maximum forest size
```{r}
ggplot(allres, aes(y = maxgrowth))+ 
  geom_boxplot() + 
  theme_minimal() + 
  labs(title = "Maximum Forest Size") +
  theme(axis.text.x = element_blank())
```

## Record the two Sobol indices (S and T)

```{r}
# "tell" the sensitivity analysis object - re-name so we keep separate
sens_C_maxgrowth <- sensitivity::tell(sens_forest_size, allres$maxgrowth)

# first-order indices (main effect without co-variance)
rownames(sens_C_maxgrowth$S) <- c("r", "K", "g", "thresh")
sens_C_maxgrowth$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
rownames(sens_C_maxgrowth$T) <- c("r", "K", "g", "thresh")
sens_C_maxgrowth$T

# graph metric against most parameter that the output is most sensitive to, using #2 for color
# create a data frame
allresp = as.data.frame(cbind(sens_forest_size$X, allres))
ggplot(allresp, aes(r, maxgrowth, col=K )) +
  geom_point() +
  labs(y="Maximum growth", 
       x="r",
       title = "Maximum forest growth (dC) by sensitivity test") +
  theme_minimal()
```

### In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about
what parameters climate change might influence)

Based on the Sobol sensitivity analysis for our forest growth model, the maximum forest growth is most sensitive to the carrying capacity followed by the linear growth rate  after canopy closure both in the first order indices and total order indices. Climate change could change this by altering the growth rate before and after canopy closure and could change the model's greatest sensitivity. 


