---
title: "assignment6_forestgrowth"
format: html
date: February 24, 2026
name: Sofia Rodas
message: false
warning: false
---

# Sensitivity Analysis of a Differential Equation
```{r}
library(tidyverse)
library(deSolve)
library(here)
library(sensitivity)
```

## Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C,
and using the following parameters
```{r}
# source our function
source(here("R/forest_growth.R"))

# parms value setting
K <- 250 
r <- 0.01
g <- 2 

forest_growth <- cbind.data.frame(r = r, K = K, g = g) 

# initial forest setting
initial_forest <- 10

# create a time sequence
tm <- seq(from = 1, to = 300)

# watch it grow
forest <- ode(y = initial_forest, times = tm, forestgrowth, parms = forest_growth)

# rename the column names
colnames(forest) <- c("time", "dC")
```

Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no
uncertainty)
```{r}
# graph the results of forest growth
ggplot(as.data.frame(forest), aes(time, dC)) +
  geom_line(color = "hotpink") + 
  theme_minimal() +
  labs(title = "Forest growth change over time",
        y = "Change in growth (kgC)")
```

Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the
estimated maximum forest size (e.g maximum of 300 years) varies with these parameters
```{r}
# set the number of parameters
np <- 2000

# Assume parameters are normally distributed
# sample K with mean of 200, standard devidation of  50
K <- rnorm(mean = 250, sd = 25, n = np)

# sample r with mean of 0.01 and standard deviation of 0.001
r <- rnorm(mean = 0.01, sd = 0.001, n = np)

# sample g with mean of 2 and standard deviation of 0.2
g <- rnorm(mean = 2, sd = 0.2, n = np)

X1 <- cbind.data.frame(r = r, K = K, g = g)

# Assume parameters are normally distributed
# sample K with mean of 200, standard devidation of  50
K2 <- rnorm(mean = 250, sd = 25, n = np)

# sample r with mean of 0.01 and standard deviation of 0.001
r2 <- rnorm(mean = 0.01, sd = 0.001, n = np)

# sample g with mean of 2 and standard deviation of 0.2
g2 <- rnorm(mean = 2, sd = 0.2, n = np)

X2 <- cbind.data.frame(r = r, K = K, g = g)

# fix any negative values and they are not meaningful
X1 <- X1 |>  
  map_df(pmax, 0.0)

X2 <- X2 |>  
  map_df(pmax, 0.0)

# create our sobel object and get sets of parameters for running the model
sens_forest_size <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# our parameter sets are
head(sens_forest_size$X)

# lets add names
colnames(sens_forest_size$X) <- c("r", "K", "g")

head(sens_forest_size$X)
```
```{r}
# turn computing our metrics into a function
compute_pop_metrics <- function(result) {
  # first add maximum population
  maxgrowth <- max(result$P)
  
  # return results
  return(list(maxgrowth = maxgrowth))
}

# Test that it works
# try it on our first parameter set, and look at when it gets to 100
# initial population
initial_forest <- 10

# run for year = 300
simtimes <-  300

# use the first parameter set (or you could just pick a test value), not the list because ode wants parameters as a list, so name each
parmtest <- list(r = sens_forest_size$X[1,"r"], 
                 K = sens_forest_size$X[1,"K"], 
                 g = sens_forest_size$X[1,"g"])

# run ode
result = ode(y = initial_forest, times = simtimes, func = forestgrowth, parms = parmtest)
# turn ode results into a labelled data frame 
result = data.frame(time=result[,1], P=result[,2])
```
Need to run this without the wrapper to see why this isn't working
### Make a wrapper function to run a Sobol test 
```{r}
# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

# lets make the threshold 90% of carrying capacity
p_wrapper <- function(r, K, g, initial_forest, simtimes, odefunc, metricfunc) {
  
  # create parameter list
  parms <- list(r = r, K = K, g = g)
  
  # run ode
  result <- ode(y = initial_forest, times = simtimes, func = odefunc, parms = parms, method="euler")
  
  # results as a data frame
  result = data.frame(time=result[,1], P=result[,2])
  # get metrics
  
  # compute metrics
  metrics <- metricfunc(as.data.frame(result))
  return(metrics)
}
```



```{r}
# test with specific r and K values (r=0.01, K=150)
# set out threshold 
p_wrapper(
  r = 0.01, K = 250, g = 2, initial_forest = 10, simtimes = 300,
 odefunc = forestgrowth, metricfunc = compute_pop_metrics
)

```

