---
title: "Assignment 7 - Two Variables ODE"
format: html
author: Sofia Rodas
date: March 10, 2026
message: false
warning: false
---

```{r}
# load necessary libraries
library(sensitivity)
library(tidyverse)
library(deSolve)
library(lhs)
library(here)
library(ggpubr)
```

Source the Lotka Voltera Model without the hunting parameter.
```{r}
source(here("R/lotvmodK.R"))
lotvmodK
```
Run the function without hunt.
```{r}
# initial conditions
currpop <- c(prey = 40, pred = 1)

# time points to see results
days <- seq(from = 1, to = 1000, by = 1)


# set parameters
pars <- c(rprey = 0.95, 
          alpha = 0.01, 
          eff = 0.6, 
          pmort = 0.4, 
          K = 2000, 
          pred_thresh = 50)

# run the model
res <- ode(func = lotvmodK, 
           y = currpop, 
           times = days, 
           parms = pars)
```

```{r}
# graph the results
head(res)

# rearrange for easy plotting
resl <- as.data.frame(res) |> 
  pivot_longer(-time, names_to = "animal", values_to = "pop")

p1 <- ggplot(resl, 
             aes(time, 
                 pop, 
                 col = animal)) +
  geom_line()

p1
```
Source the Lotka Voltera Model with the hunting parameter.
```{r}
source(here("R/lotvmodK_hunt.R"))
lotvmodK_hunt
```

Run the function with ODE. 
```{r}
# initial conditions
currpop <- c(prey = 40, pred = 1)

# time points to see results
days <- seq(from = 1, to = 1000, by = 1)

# set parameters
pars <- c(rprey = 0.95, 
          alpha = 0.01, 
          eff = 0.6, 
          pmort = 0.4, 
          K = 2000,
          rhunt = 0.7,
          prey_thresh = 50)

problem <- lotvmodK_hunt(t = 0, pop = currpop, pars = pars)
problem

# run the model
res <- ode(func = lotvmodK_hunt, 
           y = currpop, 
           times = days, 
           parms = pars,
           method = "euler")
```

```{r}
# graph the results
head(res)

# rearrange for easy plotting
resl <- as.data.frame(res) |> 
  pivot_longer(-time, names_to = "animal", values_to = "pop")

p1 <- ggplot(resl, 
             aes(time, 
                 pop, 
                 col = animal)) +
  geom_line()

p1
```


# Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the estimated populations vary with these parameters
```{r}
# set the number of parameters
np <- 200

# Assume parameters are normally distributed
K <- rnorm(mean = 2000, sd = 25, n = np)
rprey <- rnorm(mean = 0.95, sd = 0.05, n = np)
alpha <- rnorm(mean = 0.01, sd = 0.001, n = np)
eff <- rnorm(mean = 0.6, sd = 0.1, n = np)
pmort <- rnorm(mean = 0.4, sd = 0.1, n = np)
rhunt <- rnorm(mean = 0.5, sd = 0.1, n = np)
prey_thresh <- rnorm(mean = 150, sd = 20, n = np)

X1 <- cbind.data.frame(K=K,
                       rprey = rprey,
                       alpha = alpha,
                       eff = eff,
                       pmort = pmort,
                       rhunt = rhunt,
                       prey_thresh = prey_thresh)

# set the number of parameters
np <- 200

# Assume parameters are normally distributed
K2 <- rnorm(mean = 2000, sd = 25, n = np)
rprey2 <- rnorm(mean = 0.95, sd = 0.05, n = np)
alpha2 <- rnorm(mean = 0.01, sd = 0.001, n = np)
eff2 <- rnorm(mean = 0.6, sd = 0.1, n = np)
pmort2 <- rnorm(mean = 0.4, sd = 0.1, n = np)
rhunt2 <- rnorm(mean = 0.5, sd = 0.1, n = np)
prey_thresh2 <- rnorm(mean = 150, sd = 20, n = np)

X2 <- cbind.data.frame(K=K2,
                       rprey = rprey2,
                       alpha = alpha2,
                       eff = eff2,
                       pmort = pmort2,
                       rhunt = rhunt2,
                       prey_thresh = prey_thresh2)


# fix any negative values 
X1 <- X1 |>  
  map_df(pmax, 0.0)

X2 <- X2 |>  
  map_df(pmax, 0.0)

# create our sobel object and get sets of parameters for running the model
sens <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# our parameter sets are
head(sens$X)

# lets add names
colnames(sens$X) <- c("K", 
                      "rprey",
                       "alpha",
                       "eff",
                       "pmort",
                       "rhunt",
                       "prey_thresh")

head(sens$X)
```
Since we are only interested in the maximum amount of growth for our animal populations. Code the metric output function and ensure that it works with the first row of the parameters. 
```{r}
# turn computing our metrics into a function
compute_pop_metrics <- function(result) {
  # first add maximum population
  maxrprey <- max(result$prey)
  maxrpred <- max(result$pred)

  # return results
  return(list(maxrprey = maxrprey, maxrpred = maxrpred))
}

# test that the metrics computing function works
# initial population
currpop <- c(prey = 40, pred = 1)

# time points to see results
days <- seq(from = 1, to = 1000, by = 1)

# use the first parameter set for the ode function 
parmtest <- list(K = sens$X[1,"K"], 
                 rprey = sens$X[1,"rprey"], 
                 alpha = sens$X[1,"alpha"],
                 eff = sens$X[1, "eff"],
                 pmort = sens$X[1, "pmort"],
                 rhunt = sens$X[1, "rhunt"],
                 prey_thresh = sens$X[1, "prey_thresh"])

# run ode
# run the model
res <- ode(func = lotvmodK_hunt, 
           y = currpop, 
           times = days, 
           parms = parmtest)

result <- data.frame(time=res[,1], prey=res[,2], pred=res[,3])
compute_pop_metrics(result)

# turn ode results into a labelled data frame 
# result <- data.frame(time=result[,1], prey=result[,2], pred = result[,3])
```

### Make a wrapper function to run a Sobol test 
```{r}
# define a wrapper function to do everything we need - run solver and compute metrics - and send back results 
p_wrapper <- function(K, 
                      rprey, 
                      alpha, 
                      eff, 
                      pmort, 
                      rhunt, 
                      prey_thresh, 
                      currpop, 
                      simtimes, 
                      odefunc, 
                      metricfunc) {
  
  # create parameter list
  parms <- list(K = K,
                rprey = rprey,
                alpha = alpha,
                eff = eff,
                pmort = pmort,
                rhunt = rhunt,
                prey_thresh = prey_thresh)
  
  # run the model
res <- ode(func = odefunc, 
           y = currpop, 
           times = simtimes, 
           parms = parms)
  
  # results as a data frame
  result <- data.frame(time=res[,1], prey=res[,2], pred=res[,3])
  
  # compute metrics
   metrics <- compute_pop_metrics(as.data.frame(result))
   
  return(metrics)
}
```

Test the wrapper function with specific values for .
```{r}
# 
# set the threshold 
result <- p_wrapper(K = 2000, 
                    rprey = 0.95, 
                    alpha = 0.01, 
                    eff = 0.6, 
                    pmort = 0.4, 
                    rhunt = 0.1, 
                    prey_thresh = 150, 
                    currpop = currpop, 
                    simtimes = days, 
                    odefunc = lotvmodK_hunt, 
                    metricfunc = compute_pop_metrics
          )

result
```
Try the wrapper function with the first row of the varied parameters.  
```{r}
# try with first row of our Sobol parameters
p_wrapper(K = sens$X[1,"K"], 
          rprey = sens$X[1,"rprey"], 
          alpha = sens$X[1,"alpha"],
          eff = sens$X[1, "eff"],
          pmort = sens$X[1, "pmort"],
          rhunt = sens$X[1, "rhunt"],
          prey_thresh = sens$X[1, "prey_thresh"],
          currpop = currpop,
          simtimes = days,
          odefunc = lotvmodK_hunt, 
          metricfunc = compute_pop_metrics)
```
Run the wrapper function with the entirety of the varied parameters. 
```{r}
#| results: hide
#| warning: false
#| message: false

allresults <- as.data.frame(sens$X) |>  pmap(p_wrapper,
                                             currpop = currpop,
                                             simtimes = days,
                                             odefunc = lotvmodK_hunt, 
                                            metricfunc = compute_pop_metrics)

allresults <- bind_rows(allresults)

allresults
```
## Record the two Sobol indices (S and T)

```{r}
# "tell" the sensitivity analysis object - re-name so we keep separate
sens_maxgrowthr <- sensitivity::tell(sens, allresults$maxrprey)

# first-order indices (main effect without co-variance)
rownames(sens_maxgrowthr$S) <- c("K", 
                                "rprey",
                                "alpha",
                                 "eff",
                                "pmort",
                                "rhunt",
                                "prey_thresh")
sens_maxgrowthr$S

# total sensitivity index -note that this partitions the output variance - so values sum to 1
rownames(sens_maxgrowthr$T) <- c("K", 
                                "rprey",
                                "alpha",
                                 "eff",
                                "pmort",
                                "rhunt",
                                "prey_thresh")
sens_maxgrowthr$T

# graph metric against most parameter that the output is most sensitive to, using #2 for color
# create a data frame
allresp <- as.data.frame(cbind(sens$X, allresults))

ggplot(allresp, aes(rhunt, maxrprey, col=eff )) +
  geom_point() +
  theme_minimal()
```
# Metric to stability. Prey at the threshold level. 
```{r}

```

